### 2019-10-20 tutorial/03_interactive_rmd/election-interactive.Rmd
initial tutorail

---
layout: page
title: "Computational Document"
subtitle: "인터랙티브 문서"
author:
    name: 이광춘 (KPMG)
    affilates: 데이터 과학자
date: "`r Sys.Date()`"
always_allow_html: yes
output:
  html_document: 
    toc: yes
    toc_float: true
    highlight: tango
    code_folding: show
    number_section: true
    self_contained: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE,
                      comment="", digits = 3, tidy = FALSE, prompt = FALSE, fig.align = 'center')

library(reticulate)
use_condaenv("anaconda3")
# reticulate::repl_python()
```

[RStudio Trainer Program - "Tidyverse: Visualization - Lecture Note"](https://statkclee.github.io/rstudio-trainer/viz-lecture-note.html)