### 2020-02-26 cd-dashboard.Rmd
start dashboard

---
layout: page
title: "데이터 과학을 위한 저작도구: Computational Documents"
subtitle: "대쉬보드(Dashboard)"
author:
    name: "[Tidyverse Korea](https://www.facebook.com/groups/tidyverse/)"
date: "`r Sys.Date()`"
output:
  html_document: 
    toc: yes
    toc_float: true
    highlight: tango
    code_folding: show
    number_section: true
    self_contained: true
editor_options: 
  chunk_output_type: console
---

```{r setup2, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE,
                      comment="", digits = 3, tidy = FALSE, prompt = FALSE, fig.align = 'center')

library(reticulate)
# use_condaenv("anaconda3")
# reticulate::repl_python()
```

# 대쉬보드 [^dashboard] {#dashboard}

[^dashboard]: [Marlorie Hughes (2019. 12. 3), "Dashboarding Like a Boss", DC R Conference](https://www.youtube.com/watch?v=yott4quKN6s)

**A 12 Step Program to Quick Emailing Plots & Csvs**

1. Basic Layout
    - [`flexdashboard`](https://rmarkdown.rstudio.com/flexdashboard/)
1. Printing Tables, Descriptive Statistics, Model Summaries
    - `printr`, `summarytools::dfSummary`, `stargazer::stargazer`
    - [list of statistic codes](https://rdrr.io/cran/stargazer/man/stargazer_stat_code_list.html)
1. Interactive Datasets
    - `DT`
1. Interactive Plots
    - `highcharter`
    -  지도(`leaflet`), 시계열(`dygraphs`), `ggplots`(`plotly`)
1. Exporting
    - Datasets with `DT`
    - Highcharter Plots
1. CSS basics
1. Branding: Logos, Fonts, Colors
1. Header/Footer
1. Hosting
1. Automate-ish (Render + System())
1. Replicate (Render with arguments)
1. Automate or App it (cron, shiny runtime)