### 2020-02-28 cd-corona.Rmd
corona flexdashboard layout

---
title: "코로나19"
subtitle: "대쉬보드"
output: 
  flexdashboard::flex_dashboard:
    theme: readable
    orientation: rows
    vertical_layout: scroll
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
```

전세계 {data-icon="fa-globe"}
===================================

Row {data-width=150} 
--------------------------------------------

### 확진자수

```{r}
infected <- 100
valueBox(infected, icon = "fa-procedures")
```

### 사망자수

```{r}
death <- 20
valueBox(death, icon = "fa-skull")
```

### 회복자수

```{r}
recovered <- 30
valueBox(recovered, 
         icon = "fa-walking",
         color = ifelse(recovered > 10, "warning", "primary"))
```

Column {data-height=350}
-------------------------------------

### Chart 1

### Chart 2

한국 {data-icon="fa-map"}
===================================