### 2020-02-28 cd-corona.Rmd
added valuebox, leaflet, racing barchart

---
title: "코로나19"
subtitle: "대쉬보드"
output: 
  flexdashboard::flex_dashboard:
    theme: readable
    orientation: rows
    vertical_layout: scroll
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
```

전세계 {data-icon="fa-globe"}
===================================

Row {data-width=150} 
--------------------------------------------

```{r ingest-corona}
library(coronavirus)
data('coronavirus')

coronavirus <- coronavirus %>% 
  janitor::clean_names()

cv_stat <- coronavirus %>% 
  group_by(type) %>% 
  summarise(합계 = sum(cases))

```

### 확진자수

```{r infected}
infected <- cv_stat %>% filter(type=='confirmed') %>% pull
valueBox(scales::comma(infected), 
         icon = "fa-procedures",
         color = "warning")
```

### 사망자수

```{r death}
death <- cv_stat %>% filter(type=='death') %>% pull
valueBox(scales::comma(death), 
         icon = "fa-skull",
         color = "danger")
```

### 회복자수

```{r recovered}
recovered <- cv_stat %>% filter(type=='recovered') %>% pull
valueBox(scales::comma(recovered), 
         icon = "fa-walking",
         color = "success")
```

Column {data-height=350}
-------------------------------------

### Chart 1

참고: https://evamaerey.github.io/little_flipbooks_library/racing_bars/racing_barcharts.html

```{r time-series-plot}

coronavirus %>% 
  filter(type == 'confirmed') %>% 
  group_by(country_region, date) %>% 
  summarise(cases = sum(cases)) %>% 
  arrange(date, -cases) %>% 
  ungroup() %>%
  filter(str_detect(country_region, "Mainland|Korea|Japan")) %>%
  ggplot(aes(x=country_region, y=cases, fill=country_region)) +
    geom_col(show.legend = TRUE) +
    coord_flip() +
    facet_wrap(~date)

# corona_by_country %>% 
#   facet_null() %>% 
#   aes(group = country_region) +  
#   gganimate::transition_time(date)

```

### Chart 2

```{r corona-leaflet}
library(leaflet)
library(janitor)

coronavirus %>% 
  mutate(geo_name = glue::glue("{country_region} {province_state}")) %>% 
  select(-country_region, -province_state) %>% 
  group_by(geo_name, lat, long, type) %>%
    summarise(cases = sum(cases, na.rm = TRUE)) %>% 
    spread(type, cases, fill=0) %>% 
    leaflet() %>%
    addTiles() %>% 
    addProviderTiles(providers$OpenStreetMap) %>% 
    addMarkers(lng=~long, lat=~lat, clusterOptions = markerClusterOptions(),
                     popup = ~ as.character(paste0("<strong> 코로나19 현황: ", geo_name, "</strong><br>",
    "-----------------------------------------------------------<br>",
                                                   "&middot; 감염: ", scales::comma(confirmed), "<br>",
                                                   "&middot; 사망: ", scales::comma(death), "<br>",
                                                   "&middot; 회복: ", scales::comma(recovered), "<br>")))
```

한국 {data-icon="fa-map"}
===================================