### 2020-05-27 tutorial/fonts/documents.Rmd

---
title: "R마크다운 글꼴(font)"
subtitle: "문서 한글 글꼴 적용"
author:
    name: "[Tidyverse Korea](https://www.facebook.com/groups/tidyverse/)"
date: "`r Sys.Date()`"
output:
  html_document: 
    toc: yes
    toc_float: true
    highlight: tango
    code_folding: show
    number_section: true
    self_contained: true
editor_options: 
  chunk_output_type: console
---

<style>
body {
  font-family: Nanum Pen Script;
}
</style>

```{r setup2, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE,
                      comment="", digits = 3, tidy = FALSE, prompt = FALSE, fig.align = 'center')
```

# 붓꽃 데이터

붓꽃 데이터는 사랑입니다.