### 2020-09-23 datayanolja.Rmd
failed powerpoint pptx

---
title: My presentation
output: 
  powerpoint_presentation:
    reference_doc: data/slide guides_C_template.pptx
    slide_level: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Section Header

Hi

## Section header

Second