### 2020-10-25 datayanolja_template.Rmd
added ppt to xaringan conversion slideshow

---
title: "Multi-image slide layouts in PowerPoint via R Markdown"
author: "Matt Herman"
output: 
  powerpoint_presentation:
    reference_doc: "data/template-no-title.pptx"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  dev = "svg",
  fig.width = 12,
  fig.height = 7
  )
```

```{r}
library(tidyverse)
library(gapminder)
library(glue)
library(scales)
library(gridExtra)
library(patchwork)
countries <- c("Malawi", "Zambia", "Tanzania", "Kenya", "Mozambique") 
east_africa <- gapminder %>% 
  filter(country %in% countries) %>%
  mutate(country = fct_reorder2(country, year, lifeExp))
bar <- east_africa %>% 
  filter(year == max(year)) %>% 
  ggplot(aes(x = country, y = lifeExp, fill = country)) +
  geom_col(width = 0.75, alpha = 0.9) +
  geom_text(
    aes(label = number(lifeExp, 0.1)),
    position = position_stack(vjust = 0.5),
    color = "white",
    fontface = "bold",
    size = 5
    ) +
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(expand = expand_scale(0.01, 0.05)) +
  labs(y = "Life Expecancy (years)") +
  theme_minimal(base_size = 16) + 
  theme(
    legend.position = "none",
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 10),
    axis.title.y.left = element_text(margin = margin(r = 10)),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    )
tab <- east_africa %>% 
  filter(year == max(year)) %>% 
  transmute(
    Country = country, 
    Population = comma(pop),
    `GDP per capita` = dollar(gdpPercap, 1),
    `Life Expectancy` = number(lifeExp, 0.1),
    ) %>% 
  arrange(Country) %>% 
  tableGrob(theme = ttheme_minimal(), rows = NULL)
line <- east_africa %>%
  ggplot(aes(x = year, y = lifeExp, color = country)) +
  geom_line(lwd = 1.25,
            key_glyph = "timeseries") +  # for those cute glyphs in the legend
  scale_color_brewer(palette = "Dark2") +
  labs(y = "Life Expecancy (years)") +
  theme_minimal(base_size = 16) +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 10),
    axis.title.y.left = element_text(margin = margin(r = 10)),
    panel.grid.minor = element_blank(),
    plot.margin = margin(t = 30)
  )
layout <- (bar + tab) / line
layout +
  plot_annotation(
  title = "Life Expectancy of Selected Countries in East Africa",
  caption = "Source: gapminder: Data from Gapminder
             github.com/jennybc/gapminder
             gapminder.org/data/",
  theme = theme(plot.title = element_text(size = 20, hjust = 0.5, face = "bold"))
  )
```