### 2020-10-27 tutorial/slideshows/xaringanthemer.Rmd

---
title: "Tidyverse Slideshows"
subtitle: "hocolate Theme"  
author: 
  - "이광춘"
  - "Yongfu Liao"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      beforeInit: "macros.js"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
    seal: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#1381B0",
  secondary_color = "#FF961C",
  inverse_header_color = "#FFFFFF",
  header_font_google = google_font("Nanum Pen Script"),
  text_font_google   = google_font("Noto Sans KR", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

class: title-slide

.bg-text[
# Tidyverse Slideshows
### Chocolate Theme
<hr />
`r format(Sys.Date(), '%B %e, %Y')`  
Yongfu Liao
이광춘
]

---
## Table of Contents

- Get Started
    - Hello World
    - Tidyverse Korea

---
class: center, middle

# Get Started

---
## Hello World

Install the **xaringan** package from [Github](https://github.com/yihui/xaringan):

```{r eval=FALSE, tidy=FALSE}
devtools::install_github("yihui/xaringan")
```

--

You are recommended to use the [RStudio IDE](https://www.rstudio.com/products/rstudio/), but you do not have to.

- Create a new R Markdown document from the menu `File -> New File -> R Markdown -> From Template -> Ninja Presentation`;<sup>1</sup>

.center[
<img src="fig/xaringan-hex.png" width="17%" />  
]

.footnote[
[1] 中文用户请看[这份教程](http://slides.yihui.name/xaringan/zh-CN.html)
[2] See [#2](https://github.com/yihui/xaringan/issues/2) if you do not see the template or addin in RStudio.
]

---
## Tidyverse Korea

.pull-left[
```{r matching-ggplot, eval = FALSE}
library(ggplot2)
ggplot(diamonds) +
  aes(cut, fill = cut) +
  geom_bar(show.legend = FALSE) +
  labs(
    x = "Cut",
    y = "Count",
    title = "A Fancy diamonds Plot"
  ) +
  theme_xaringan(background_color = "#FFFFFF") +
  scale_xaringan_fill_discrete()
```
]

.pull-right[
```{r matching-ggplot-plot, echo = FALSE, eval = TRUE}
ggplot(diamonds) +
  aes(cut, fill = cut) +
  geom_bar(show.legend = FALSE) +
  labs(
    x = "Cut",
    y = "Count",
    title = "A Fancy diamonds Plot"
  ) +
  theme_xaringan(background_color = "#FFFFFF") +
  scale_xaringan_fill_discrete()
```
]

---
class: center, middle

# Thanks!

Slides created via the R package [**xaringan**](https://github.com/yihui/xaringan).

The chakra comes from [remark.js](https://remarkjs.com), [**knitr**](http://yihui.name/knitr), and [R Markdown](https://rmarkdown.rstudio.com).