### 2021-01-13 cd-static-website-blog.Rmd
started website blog

---
layout: page
title: "데이터 과학을 위한 저작도구: Computational Documents"
subtitle: "정적 웹사이트 + 블로그"
author:
- name: "이광춘"
  affiliation: "[Tidyverse Korea](https://www.facebook.com/groups/tidyverse/)"
tags: ["데이터 과학", "Data Science", "데이터 사이언스", "distill", "웹사이트", "과학기술", "블로그"]
date: "`r Sys.Date()`"
output:
  html_document: 
    include:
      after_body: footer.html
      before_body: header.html
    theme: default
    toc: yes
    toc_depth: 2
    toc_float: true
    highlight: tango
    code_folding: show
    number_section: true
    self_contained: yes
editor_options: 
  chunk_output_type: console
bibliography: bibliography-cd.bib
csl: biomed-central.csl
---

```{r setup2, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE,
                      comment="", digits = 3, tidy = FALSE, prompt = FALSE, fig.align = 'center')

library(reticulate)
```

# 웹사이트 아키텍처 [@alison2020blog] {#website-architecture}

![](fig/distill-website-blog.png)