### 2021-03-25 cd-resume-automation.Rmd
start resume automation

---
layout: page
title: "데이터 과학을 위한 저작도구: Computational Documents"
subtitle: "이력서 자동화: `css` + `googlesheets` + 웹사이트"
author:
  - name: "이광춘"
    affiliation: "[Tidyverse Korea](https://www.facebook.com/groups/tidyverse/)"
date: "`r Sys.Date()`"
tags: ["데이터 과학", "Data Science", "데이터 사이언스", "resume", "CV", "이력서", "pagedown"]
output:
  html_document: 
    include:
      after_body: footer.html
      before_body: header.html
    theme: default
    toc: yes
    toc_depth: 2
    toc_float: true
    highlight: tango
    code_folding: show
    number_section: true
    self_contained: true
bibliography: bibliography-cd.bib
csl: biomed-central.csl
urlcolor: blue
linkcolor: blue
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE,
                      comment="", digits = 3, tidy = FALSE, prompt = FALSE, fig.align = 'center')

library(pdftools)
library(tidyverse)
library(magick)
```

![](fig/pagedown-resume.png)

#