### 2021-07-13 business_card/R_user_group.Rmd
added business card

---
phone: "+82 010-xxxx-xxxx"
email: "victor@r2bit.com"
url: www.r2bit.com
address: |
  경기도 성남시 분당구 판교
logo: "https://raw.githubusercontent.com/statkclee/art/gh-pages/fig/R_user_group.png"
person: 
  - name: 이광춘
    title: R 사용자회, 대표
    repeat: 12
googlefonts: "sunflower" # feel free to try other fonts at https://fonts.google.com
paperwidth: 8.5in
paperheight: 11in
cols: 4
rows: 3
output: pagedown::business_card
---

<!-- if you prefer black text on white background, set eval=TRUE on the chunk below -->

```{css, eval=TRUE}
.wrapper {
  color: black;
  background-color: skyblue;
  border: 1px dotted black;
}
.coordinates {
  color: #333;
}
```