### 2021-07-25 ds-css-template.Rmd
start xaringan template

---
layout: page
title: "데이터 과학을 위한 저작도구: Computational Documents"
subtitle: "`css` 템플릿: 글꼴, 스타일, ..."
author:
  - name: "이광춘"
    affiliation: "[Tidyverse Korea](https://www.facebook.com/groups/tidyverse/)"
date: "`r Sys.Date()`"
tags: ["데이터 과학", "Data Science", "데이터 사이언스", "tidymodels", "parsnip", "로지스틱 회귀모형", "tidyverse", "팽귄", "penguin", "CSS", "Template"]
output:
  html_document: 
    include:
      after_body: footer.html
      before_body: header.html
    theme: default
    toc: yes
    toc_depth: 2
    toc_float: true
    highlight: tango
    code_folding: show
    number_section: true
    self_contained: true
bibliography: bibliography-cd.bib
csl: biomed-central.csl
urlcolor: blue
linkcolor: blue
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE,
                      comment="", digits = 3, tidy = FALSE, prompt = FALSE, fig.align = 'center')

library(pdftools)
library(tidyverse)
library(magick)

```

![](fig/design_specs.png){width="357"}

# PPT 스타일 [\^xaringan-tutorial] {#xaringan-style}

```{r ppt-style}
library(tidyverse)
library(magick)
library(sysfonts)
font_add_google("Noto Sans KR")
# library(extrafont)
# loadfonts()

template_img <- image_blank(width = 400, height = 500, color = "darkgray", pseudo_image = "", defines = NULL)

template_img %>% 
  image_annotate("Typography", location = "+30+30", size = 20, font = "NanumGothic", color = "black") %>% 
  image_annotate("Aa가", location = "+30+50", size = 30, font = "Noto Sans KR", color = "black")

```